---
title: "R Cheat Sheet"
output: html_notebook
---

# Vectors
Main building block in R.
```{r}
c(1, 2, 3, 4)
```

```{r}
is.vector(c(1, 2, 3, 4))
```

#### Functions for vectors
`sum()`, `mean()`, `length()`, `rev()`, `unique()`, `table()`, `sort()`
```{r}
# Example
length(c(1, 2, 3, 4))
```

#### Functions that make vectors
`:`, `seq()`, `rep()`
```{r}
# Examples
1:4
seq(0, 10, by = 2)
rep(1:3, times = 4)
rep(1:3, each = 4)
```

#### Operators on vectors
`+`, `*`, `/`, `-`, and `^` for powers.
Integer division `%/% and mod %%.
Integer division tells you how many times a second number divides into the first to the nearest whole number. Mod tells you what's left over after this division.
These operators work on vectors of any size. In one command we can divide all elements.
```{r}
# Example
1:5 / 2
```

#### Operators that make logical vectors
`==`, `!=` `<`, `>`, `<=`, `>=`, `%in%`, `&&`, `||`
```{r}
# Examples
"b" %in% c("a", "b", "c")
TRUE && FALSE
TRUE || FALSE
# get help on relational operators by using 
?`!=`
```

#### Named vectors
You can add names to a vector by using the function `names` and an assignment.
```{r}
x <- c(1, 2, 3, 4, 5, 6, 7)
names(x) <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
names(x)
x
```

#### Subsetting vectors
```{r}
x[3]
x[1:3]
x[c(1, 4, 7)]
x[-3]
```
For named vectors, you can select by name.
```{r}
x["Sun"]
```

Subset with a logical vector.
```{r}
greater_than_4 <- x > 4
greater_than_4
```
Now using it to subset.
```{r}
x[greater_than_4]
```
In one step we could use.
```{r}
x[x > 4]
```

# Data Frames
Construct a data frame from vectors.
```{r}
person <- c("David", "Susan", "Molly")
height <- c(1.7, 1.5, 1.2)
shoe_size <- c(8, 7, 6)
information_collected <- data.frame(person, height, shoe_size)
information_collected
```

```{r}
information_collected$height
```
If we want to add an extra vector to our data frame, we can assign a new name to that data frame directly.
```{r}
information_collected$question <- c(1L, 2L, 3L)
information_collected
```
Deleting a vector (column) from a data frame can be done by setting to null.
```{r}
information_collected$question <- NULL
information_collected
```
The function `str` gives you respresentation of what is in an object.
```{r}
str(information_collected)
```
You can use `View(data frame)` to view your data frame. You can also click on the item in the `Global Environment` to also do this.

#### Selecting data-frames
You can select data frames by number.
To select a row:
```{r}
information_collected[1, ]
```

To select a column:
```{r}
information_collected[, 1]
```

Or select a specfic item:
```{r}
information_collected
information_collected[2, 2]
```

#### Sorting data-frames
By creating this:
```{r}
height_order <- order(information_collected$height)
```
We can then use it to select:
```{r}
information_collected[height_order, ]
```

# Vectors
A list is a special data type in R that acts as a container. Unlike Vectors, the contents of a list are not restricted to a single type.

A list can contain any type of R onject, including another list.

Create lists using `list` or change another object to a list using `as.list`.
```{r}
y <- list("item a" = 1, "item b" = "b", "item c" = TRUE, "item d" = c(1, 2, 3))
y
```

```{r}
y[2]
y[[2]]
```

```{r}
y$`item b`
```


